\documentclass[11pt,addpoints]{exam}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}
\usepackage{mhchem}

\title{Baseline Quiz}
\author{Mark Peever\\ \texttt{mpeever@gmail.com}}
\date{August 22, 2025}

\begin{document}
%\maketitle

\pointsinrightmargin
%\marginpointname{ \points}
\bonuspointpoints{bonus point}{bonus points}

\begin{center}
\fbox{\fbox{\parbox{5.5in}{\centering
Answer the questions on a separate paper.
}}}
\end{center}
\vspace{0.1in}
\makebox[\textwidth]{Name:\enspace\hrulefill}
\vspace{0.2in}

\begin{questions}

\question
What are the names of the following elements?

\begin{parts}
\part[1] Na
\part[1] Pb
\part[1] N
\part[1] K
\end{parts}
\vspace{.2in}

\question
Reduce the following fractions:

\begin{parts}
\part[1] $\frac{120}{400}$

\part[1] $\frac{256 \pi}{32}$

\part[1] $\frac{x^2-1}{x + 1}$

%\part[1] $\frac{2x^2+4x+2}{x^2-1}$
\end{parts}
\vspace{.2in}


\question
Multiply the following fractions:

\begin{parts}
\part[1]
$\frac{3 \pi}{4} \times \frac{1}{2 \pi}$

\part[1]
$\frac{256 \pi}{32} \times 12$

%\part[1]
%$\frac{x^2-1}{x + 1} \times \frac{2}{x - 1}$

\part[1]
$\frac{2x^2+4x+2}{x^2-1} \times \frac{3x}{6}$

\end{parts}
\vspace{.2in}

\question
Convert the following measurements:

\begin{parts}
%\part[1]
%how many inches in 3 yards?

\part[1]
how many tablespoons in one gallon?

%\part[1]
%how many pounds in 373 ounces?

\part[1]
how many kilograms in $1,000,000,000$ grams?

\part[1]
how many centimeters in 375 meters?
\end{parts}
\vspace{.2in}


\question[1]
It is a truth universally acknowledged, that a single man in possession of a good fortune, must be in want of \fillin[a wife].
\vspace{.2in}

\pagebreak

\question[2] What is $ 12.0 \times 3.21 $? Choose the most correct answer:
\begin{choices}
\choice 39
\choice 38.5
\choice 38.52
\choice 38.520
\choice 38.5200
\end{choices}
\vspace{.2in}

\question[2] What is $ 12.0 + 3.21 $? Choose the most correct answer:
\begin{choices}
\choice 15
\choice 15.2
\choice 15.21
\choice 15.210
\choice 15.2100
\end{choices}
\vspace{.2in}

\question[1] How much wood would a woodchuck chuck if a woodchuck could chuck wood?
\vspace{.2in}

\question[1] The boiling point of water in Seattle is $100^{\circ}$C ($212^{\circ}$F), the boiling point of water in Moscow is $97^{\circ}$C ($206^{\circ}$F). Why is it lower in Moscow?
\vspace{.2in}

\question Express the following in scientific notation:
\begin{parts}
\part[1] $0.00000674$
\part[1] $2048$
\end{parts}
\vspace{.2in}

\question Give definitions for the following:
\begin{parts}
\part[1] heat
\part[1] electrolyte
\end{parts}
\vspace{.2in}

\question Solve the following equations for the named variable:
\begin{parts}
\part[1] $12 = \frac{2}{5} x$ (solve for $x$)

\part[1] $3t - 17 = 22 $ (solve for $t$)

%\part[1] $\frac{y^2-1}{y + 1} = 0$ (solve for $y$)

\part[1] $5 \beta + 12 = 2 \beta$ (solve for $\beta$)

\part[1] $\frac{2\theta - 1}{2\theta + 1} = 3$ (solve for $\theta$)

\end{parts}
\pagebreak
%\vspace{.2in}

\question[5] What happens when you leave an ice tray full of water the deep freezer for several days?
Choose all the correct answers.

\begin{choices}
\choice the water gets warmer
\choice the molecules in the water slow down
\choice the water undergoes a physical change from liquid to gas
\choice the water changes phase from liquid to solid
\choice the water undergoes a chemical change from liquid to solid
\end{choices}
\vspace{.2in}

\question[1] 
What is the fastest known time for the Kessel Run?

\begin{choices}
\choice 72 hours
\choice 12 parsecs
\choice 14 petabytes
\choice 3 weeks
\choice 123 kiloseconds
\end{choices}
\vspace{.2in}

\question[5]
Which of the following is an example of a chemical reaction?

\begin{choices}
\choice a piece of wood burns in a bonfire
\choice a piece of steak is fed into a meat grinder
\choice one gram of table salt is dissolved into a glass of water
\choice a pan of water is brought to a boil, and left boiling until it is dry
\choice vinegar and baking soda are combined in a cup and they produce foam
\end{choices}
\vspace{.2in}

\question
Balance the following chemical equations:
\begin{parts}
\part[1] \ce{NaOH + HCl ->  H2O + NaCl } 

\part[1] \ce{CH4 (g) + O2 (g)  -> CO2 (g) + H2O (g)}	

% \part[1] \ce{LiOH + H2SO4 -> Li2SO4 + H2O}

\part[1] \ce{C3H8 (g) + O2 (g) -> CO2 (g) + H2O (g)}
\end{parts}
\pagebreak

\question[1] What is $2x^2 + 13x + 123$ if $ x = 176230 $?

\begin{choices}
\choice a whole lot
\choice a really big number
\choice something really huge
\choice a whole number larger than 176353
\choice all of the above
\end{choices}
\vspace{.2in}

\question[1] You go to the store and ask for one mole of eggs. How many eggs are you asking for?
\vspace{.2in}

\question[1] How many protons are in the nucleus of an Oxygen (\ce{O}) atom?
\vspace{.2in}

\question
Classify the following compounds as covalent, ionic, or metallic:
\begin{parts}
\part[1] \ce{H2O}
\part[1] \ce{CaCO3}
\part[1] \ce{H2SO4}
\part[1] \ce{H2O2}
%\part[1] \ce{(CH3O)2SO2}
\part[1] \ce{C3H8}
\end{parts}
\vspace{.2in}

\bonusquestion[3]
(Bonus) Three men are called ``king of kings" in Scripture, who are they?

\end{questions}

\vspace{2in}

\begin{center}
This exam has \numquestions\ questions for a total of \numpoints\ points and \numbonuspoints\ bonus points.
\end{center}

\begin{center}
\gradetable[v][questions]
\end{center}

\end{document}  