\documentclass[11pt,addpoints]{exam}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}
\usepackage{mhchem}
\usepackage{lewis}


\title{Module 10 Quiz}
\author{Mark Peever}
\date{February 27, 2026}							% Activate to display a given date or no date

\begin{document}
\maketitle
\printanswers
\pointsinrightmargin
%\marginpointname{ \points}

%\begin{center}
%\fbox{\fbox{\parbox{5.5in}{\centering
%$q = m c \Delta T$ \\
%$\rho = \frac{m}{V} $
%}}}
%\end{center}
\vspace{0.1in}
\makebox[\textwidth]{Name:\enspace\hrulefill}
\vspace{0.2in}

\begin{questions}
\question[5] \ce{HCl} is an acid. Which of the following happens when \ce{HCl} is dissolved in water?
 \begin{choices}
\choice \ce{HCl (aq) + H2O (l) -> H3O^{+} (aq) + Cl^{-}}
\choice \ce{HCl (aq) + H2O (l) -> H3O^{+} (aq) + Cl^{2-}}
\choice \ce{HCl (aq) + H2O (l) -> H3O^{2+} (aq) + 2Cl^{-}}
\choice \ce{HCl (aq) + H2O (l) -> H3O^{+} (aq) + OH^{-}}
\end{choices}

\begin{solution}
A) \ce{HCl (aq) + H2O (l) -> H3O^{+} (aq) + Cl^{-}}
\end{solution}

\vspace{0.2in}

\question[5] \ce{NaOH} is a base. Which of the following happens when \ce{NaOH} is dissolved in water?

 \begin{choices}
\choice \ce{NaOH (aq) -> Na^{+} (aq) + OH^{-} (aq)} 
\choice \ce{NaOH (aq) -> Na^{2+} (aq) + OH^{-} (aq)} 
\choice \ce{NaOH (aq) -> Na^{+} (aq) + OH^{2-} (aq)} 
\choice \ce{NaOH (aq) -> Na^{-} (aq) + H3O^{+} } 
\end{choices}

\begin{solution}
A) \ce{NaOH (aq) -> Na^{+} (aq) + OH^{-} (aq)} 
\end{solution}

\vspace{0.2in}

\clearpage
\question[5] \ce{NH3} is a base. Which of the following happens when \ce{NH3} is dissolved in water?

 \begin{choices}
\choice \ce{NH3 (aq) + H2O (l) -> NH4^{+} (aq) + OH^{-} (aq)}
\choice \ce{NH3 (aq) + H2O (l) -> NH4^{-} (aq) + OH^{-} (aq)}
\choice \ce{NH3 (aq) + H2O (l) -> NH4^{+} (aq) + 2OH^{-} (aq)}
\choice \ce{NH3 (aq) + H2O (l) -> NH4^{+} (aq) + H3O^{+} (aq)}
\end{choices}

\begin{solution}
A) \ce{NH3 (aq) + H2O (l) -> NH4^{+} (aq) + OH^{-} (aq)}
\end{solution}

\vspace{0.2in}


\question[5] Which of the following acid/base reactions is correct?

 \begin{choices}
\choice \ce{H2SO4 (aq) + HCl (aq) -> Cl2SO4 (aq) + H2O (l)} 
\choice \ce{HCl (aq) + NaOH (aq) -> NaCl (aq) + H2O (l)} 
\choice \ce{H3PO4 (aq) + HCl (aq) -> ClPO4^{4-} (aq) + H2O (l)} 
\choice \ce{H2CO3 (aq) + NaOH (aq) -> Na2CO3 (aq) + 2H2O (l)} 
\end{choices}

\begin{solution}
B)  \ce{HCl (aq) + NaOH (aq) -> NaCl (aq) + H2O (l)} 
\end{solution}

\vspace{0.2in}

\question[5] Which of the following acid/base reactions is correct?

 \begin{choices}
\choice \ce{H2SO4 (aq) + HCl (aq) -> Cl2SO4 (aq) + H2O (l)} 
\choice \ce{H2SO4 (aq) + NaOH (aq) -> Na2SO4 (aq) + H2O (l)} 
\choice \ce{HPO3 (aq) + HCl (aq) -> ClPO4 (aq) + H2O (l)} 
\choice \ce{H2SO4 (aq) + 2NaOH (aq) -> Na2SO4 (aq) + 2H2O (l)} 
\end{choices}

\begin{solution}
D)  \ce{H2SO4 (aq) + 2NaOH (aq) -> Na2SO4 (aq) + 2H2O (l)} 
\end{solution}

\vspace{0.2in}
\clearpage
\question[5] Choose the correct answer:

 \begin{choices}
\choice an acid is a proton donor
\choice a base is a proton receptor
\choice an electrolyte gives ions in aqueous solution
\choice all of the above
\choice none of the above
\end{choices}

\begin{solution}
D) all of the above
\end{solution}

\end{questions}

\begin{center}
This quiz has \numquestions\ questions for a total of \numpoints\ points.
\end{center}

\end{document}  